---
title: "covid 19"
author: "Bessie Luevano"
date: "Summer 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$sex <- as.factor(kRdata$sex)
kRdata$result <- as.factor(kRdata$result)
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```  

The barchart appears to show that male deaths related to Covid make up 60% of the total of deaths. Whereas females make up 40% of the deaths.



  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that 64% of dead people where males and 36% of the dead people where females. Males are twice as likely to die from Covid compared to females.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the inferential results, looking at Fisher's exact test for count data, if you are a male, you are 2.3 times more likely to die from Covid compared to females. 

## Conclusions
We used the Cat~cat helper file and we added kRdata$sex and kRdata$result to give us factors. We added the barchart which showed the amount of deaths of 100 and it appeared that male deaths were 60% and female deaths were 40%. The numerical results showed that 64% males died from Covid and 36% of females died from Covid. Through the inferential results we could conclude that males are 2.3 times more likely to die from Covid than females. Therefore, we can reject the null hypothesis which states that Covid deaths are independent of sex.
